---
title: "In_situ_probiotics"
author: "J. Meyer"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r, echo=FALSE}
library(dada2)
library(ShortRead)
library(ggplot2)
library(phyloseq)
library(vegan)
library(knitr)
library(ALDEx2)
library(CoDaSeq)
library(zCompositions)
library(igraph)
library(car)
library(grDevices)
library(propr)
library(cowplot)
library(randomcoloR)
library(dplyr)
library(reshape2)
library(tibble)
library(exactRankTests)
library(nlme)
library(data.table)
library(Rmisc)
library(indicspecies)
library(viridis)
#library(biomehorizon) 
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```

## Quality-filter the sequencing reads and create Amplicon Sequence Variant (ASV) tables with DADA2

Put unjoined R1 and R2 fastq files, with adaptors and primers previously removed with cutadapt into a directory for DADA2. Here, our forward and reverse fastq filenames have format: SAMPLENAME_R1_cut.fastq.gz and SAMPLENAME_R2_cut.fastq.gz

*****If you have samples from multiple sequencing runs, you need to determine the sequence variants for each run separately, then merge the ASV tables.
Here is the dada2 page on merging runs: https://benjjneb.github.io/dada2/bigdata_paired.html




BS1 site field trials Jan 2020 - abandoned due to covid, but got before and after first treatment; 3 sequencing runs total
```{r, echo=FALSE}
#### 1st sequencing run for BS1
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1968R_BS1"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS1968R_BS1.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1968R_BS1/seqtab.rds") 
```


```{r, echo=FALSE}
#### 2nd sequencing run for BS1
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1995_BS1"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS1995_BS1.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1995_BS1/seqtab.rds") 
```


```{r, echo=FALSE}
#### 3rd sequencing run for BS1
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2051_BS1"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2051_BS1.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2051_BS1/seqtab.rds") 
```


BS2 site field trials 2020-2021; 4 sequencing runs total
```{r, echo=FALSE}
#### 1st sequencing run for BS2
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2107_BS2"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2107_B2.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2107_BS2/seqtab.rds") 
```


```{r, echo=FALSE}
#### 2nd sequencing run for BS2
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2172_BS2"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2172_BS2.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2172_BS2/seqtab.rds")
```


```{r, echo=FALSE}
#### 3rd sequencing run for BS2
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2211_BS2"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2211_BS2.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2211_BS2/seqtab.rds")
```


```{r, echo=FALSE}
#### 4th sequencing run for BS2
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2408_BS2"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "-V4"), `[`, 1) #### NS2408 sequencing files named differently, can't use "_" as separator
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2408_BS2.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2408_BS2/seqtab.rds")
```



BS3 and MK-48 sites field trials 2021-2022; 2 sequencing runs to date
```{r, echo=FALSE}
#### 1st sequencing run for BS3 and MK48
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2661_BS3"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "-515rcbc"), `[`, 1)

# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2661_BS3.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2661_BS3/seqtab.rds")
```

```{r, echo=FALSE}
#### 2nd sequencing run for BS3 and MK48
path <- "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2696_BS3"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "-515rcbc"), `[`, 1)

# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats_NS2696_BS3.txt",sep="\t",col.names=NA)
saveRDS(seqtab, "~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2696_BS3/seqtab.rds")
```


Now that I have determined the ASV tables for all 9 sequencing runs, I can merge the ASV tables and remove chimera sequences.


```{r, echo=FALSE}
st1 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1968R_BS1/seqtab.rds")
st2 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS1995_BS1/seqtab.rds") 
st3 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2051_BS1/seqtab.rds")
st4 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2107_BS2/seqtab.rds") 
st5 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2172_BS2/seqtab.rds") 
st6 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2211_BS2/seqtab.rds") 
st7 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2408_BS2/seqtab.rds") 
st8 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2661_BS3/seqtab.rds")
st9 <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/cutadapt_NS2696_BS3/seqtab.rds")
st.all <- mergeSequenceTables(st1, st2, st3, st4, st5, st6, st7, st8, st9, repeats="sum") # You may get the message "Duplicated sample names detected in the sequence table row names." to let you know that there are duplicate names across samples - it is not an error, just a message. If you have run a sample on more than one sequencing run, the ASV counts will be added together.
#Remove chimeric sequences:
seqtab.nochim <- removeBimeraDenovo(st.all, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
sum(seqtab.nochim)/sum(st.all)
# SAVE the non-chimeric sequence variant table SO YOU DON'T HAVE TO REPEAT ALL OF THE ABOVE STEPS
saveRDS(seqtab.nochim, file="~/Documents/Github/McH1-7_Probiotics_Field_Trials/probiotics.rds")


# Combine read stats from all runs and add chimera summary - This does not combine duplicate rows from repeated samples.
stat1 <- read.table("dada_read_stats_NS1968R_BS1.txt",sep="\t",header=TRUE, row.names=1)
stat2 <- read.table("dada_read_stats_NS1995_BS1.txt",sep="\t",header=TRUE, row.names=1)
stat3 <- read.table("dada_read_stats_NS2051_BS1.txt",sep="\t",header=TRUE, row.names=1)
stat4 <- read.table("dada_read_stats_NS2107_BS2.txt",sep="\t",header=TRUE, row.names=1)
stat5 <- read.table("dada_read_stats_NS2172_BS2.txt",sep="\t",header=TRUE, row.names=1)
stat6 <- read.table("dada_read_stats_NS2211_BS2.txt",sep="\t",header=TRUE, row.names=1)
stat7 <- read.table("dada_read_stats_NS2408_BS2.txt",sep="\t",header=TRUE, row.names=1)
stat8 <- read.table("dada_read_stats_NS2661_BS3.txt",sep="\t",header=TRUE, row.names=1)
stat9 <- read.table("dada_read_stats_NS2696_BS3.txt",sep="\t",header=TRUE, row.names=1)
stats.all<-bind_rows(stat1, stat2,stat3,stat4,stat5,stat6,stat7,stat8,stat9) 
write.table(stats.all, "dada_read_stats_all.txt",sep="\t",col.names=NA)
# Track reads through the pipeline
# As a final check of our progress, we’ll look at the number of reads that made it through each step in the pipeline
rowSums(seqtab.nochim)
# need to write this out to add to dada read stats

# SAVE the non-chimeric sequence variant table SO YOU DON'T HAVE TO REPEAT ALL OF THE ABOVE STEPS
saveRDS(seqtab.nochim, file="~/Documents/Github/McH1-7_Probiotics_Field_Trials/probiotics.rds")
# RELOAD THE SAVED INFO FROM HERE (if you have closed the project):
#seqtab.nochim <- readRDS("~/Documents/Github/McH1-7_Probiotics_Field_Trials/probiotics.rds")
```

## Assign taxonomy in DADA2

Make sure the taxonomy reference database is in your working directory. Keep the database file gzipped. Adjust path name below. This step is very time consuming.

When taxonomy assignment is complete, we will use base R and phyloseq to clean up the taxonomy table. First, we will replace NAs and empty cells with the lowest taxonomy classification available. Second, we will use phyloseq to remove reads that are classified as Eukaryotes or unclassified at the domain level (ie, we are keeping only Bacteria and Archaea because that is what our primers target).

```{r, echo=FALSE}
taxa <- assignTaxonomy(seqtab.nochim, "~/Documents/Github/silva_nr99_v138.1_train_set.fa.gz", multithread=TRUE)
# FIX the NAs in the taxa table
taxon <- as.data.frame(taxa,stringsAsFactors=FALSE)
taxon$Phylum[is.na(taxon$Phylum)] <- taxon$Kingdom[is.na(taxon$Phylum)]
taxon$Class[is.na(taxon$Class)] <- taxon$Phylum[is.na(taxon$Class)]
taxon$Order[is.na(taxon$Order)] <- taxon$Class[is.na(taxon$Order)]
taxon$Family[is.na(taxon$Family)] <- taxon$Order[is.na(taxon$Family)]
taxon$Genus[is.na(taxon$Genus)] <- taxon$Family[is.na(taxon$Genus)]
write.table(taxon,"silva_taxa_table.txt",sep="\t",col.names=NA)
write.table(seqtab.nochim, "silva_otu_table.txt",sep="\t",col.names=NA)
# Create phyloseq object from otu and taxonomy tables from dada2, along with the sample metadata.
otu <- read.table("silva_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_17May.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps #33209 taxa and 592 samples
# remove chloroplasts and mitochondria and Eukaryota
get_taxa_unique(ps, "Family")
get_taxa_unique(ps, "Order")
get_taxa_unique(ps, "Kingdom")
ps <- subset_taxa(ps, Family !="Mitochondria")
ps <- subset_taxa(ps, Order !="Chloroplast")
ps <- subset_taxa(ps, Kingdom !="Eukaryota")
ps <- subset_taxa(ps, Kingdom !="NA")
get_taxa_unique(ps, "Family")
get_taxa_unique(ps, "Order") 
get_taxa_unique(ps, "Kingdom")
ps #31624 taxa and 592 samples; 594 in metadata

# Now export cleaned otu and taxa tables from phyloseq for future reference
otu = as(otu_table(ps), "matrix")
taxon = as(tax_table(ps), "matrix")
metadata = as(sample_data(ps), "matrix")
write.table(otu,"silva_nochloronomito_otu_table.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table.txt",sep="\t",col.names=NA)
# export ASV table as relative abundance
ps_ra<-transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
otu_ra = as(otu_table(ps_ra), "matrix")
write.table(otu_ra,"silva_nochloronomito_otu_table_RA.txt",sep="\t",col.names=NA)
```

Now, time to explore the data.

```{r, echo=FALSE}
ps #31624 taxa and 592 samples; 594 in metadata
get_taxa_unique(ps, "Order") #469
get_taxa_unique(ps, "Class") #195
ps5<-filter_taxa(ps, function(x) mean(x) >5, TRUE)
ntaxa(ps5) #456
ps10<-filter_taxa(ps, function(x) mean(x) >10, TRUE)
ntaxa(ps10) #258
get_taxa_unique(ps, "Genus") #2000
get_taxa_unique(ps5, "Genus") #219
get_taxa_unique(ps10, "Genus") #141

# Now export filtered otu and taxa tables from phyloseq for future reference after removing sample blanks
ps5 #456 taxa and 592 samples
ps5 = subset_samples(ps5, tag != "blank")
ps5 #456 taxa and 581 samples
otu_ps5 = as(otu_table(ps5), "matrix")
taxon_ps5 = as(tax_table(ps5), "matrix")
metadata = as(sample_data(ps5), "matrix")
write.table(otu_ps5,"silva_nochloronomito_otu_table_ps5.txt",sep="\t",col.names=NA)
write.table(taxon_ps5,"silva_nochloronomito_taxa_table_ps5.txt",sep="\t",col.names=NA)
write.table(metadata,"metadata_noblanks.txt",sep="\t",col.names=NA)
ps5_ra<-transform_sample_counts(ps5, function(OTU) OTU/sum(OTU))
otu_ps5_ra = as(otu_table(ps5_ra), "matrix")
write.table(otu_ps5_ra,"silva_nochloronomito_otu_table_ps5_RA.txt",sep="\t",col.names=NA)
```

## Perform center-log-ratio transformation on ASVs and calculate Aitchison Distance and principal components

```{r, echo=FALSE}
# use an ASV table that has been filtered to remove low-abundance ASVs, no blanks
otu <- read.table("silva_nochloronomito_otu_table_ps5.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps5.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_noblanks.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps #456 taxa and 581 samples
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu), method="CZM", label=0)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)
# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))

samples$month<-factor(samples$month,levels=c("Jan","May","July","Aug","Sept","Oct","Nov"))
samples$treatment <- factor(samples$treatment, levels = c("none","control bag","control paste","control paste and bag","probiotic bag","probiotic paste","probiotic paste and bag","resistant"))
treatcolors <- c("none"="#999999","control bag"="#56B4E9","control paste"="#0072B2","control paste and bag"="#000066","probiotic bag"="#E69F00","probiotic paste"="#D55E00","probiotic paste and bag"="#993300","resistant"="#000000")


pdf("PCA.pdf",bg ="white",width=8.5)
p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=samples$treatment,shape=samples$month))
p<-p+geom_point(size=5)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  scale_fill_manual(values=treatcolors)+
  scale_shape_manual(values=c(21,0,22,23,1,24,25))+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab, fill="Treatment",shape="Month") + coord_fixed()
dev.off()

bs<-subset_samples(ps, site!="Mk48-5")
bs  #456 taxa and 480 samples
otu_bs = as(otu_table(bs), "matrix")
taxon_bs = as(tax_table(bs), "matrix")
meta_bs = as(sample_data(bs), "matrix")
write.table(otu_bs,"silva_nochloronomito_otu_table_bs.txt",sep="\t",col.names=NA)
write.table(taxon_bs,"silva_nochloronomito_taxa_table_bs.txt",sep="\t",col.names=NA)
write.table(meta_bs,"metadata_bs.txt",sep="\t",col.names=NA)
otu <- read.table("silva_nochloronomito_otu_table_bs.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_bs.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_bs.txt",sep="\t",header=T,row.names=1)
d.czm <- cmultRepl(t(otu), method="CZM", label=0)
d.clr <- codaSeq.clr(d.czm)
E.clr <- t(d.clr)
d.pcx <- prcomp(E.clr)
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)
# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
samples$month<-factor(samples$month,levels=c("Jan","July","Aug","Oct","Nov"))
samples$treatment <- factor(samples$treatment, levels = c("none","control bag","control paste","control paste and bag","probiotic bag","probiotic paste","probiotic paste and bag","resistant"))
treatcolors <- c("none"="#999999","control bag"="#56B4E9","control paste"="#0072B2","control paste and bag"="#000066","probiotic bag"="#E69F00","probiotic paste"="#D55E00","probiotic paste and bag"="#993300","resistant"="#000000")

pdf("PCA_BS.pdf",bg ="white",width=8.5)
p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=samples$treatment,shape=samples$month))
p<-p+geom_point(size=5)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  scale_fill_manual(values=treatcolors)+
  scale_shape_manual(values=c(21,22,23,24,25))+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab, fill="Treatment",shape="Month") + coord_fixed()
dev.off()


## Use phyloseq/vegan to perform ANOSIM/PERMANOVA using Aitchison distance

# create a phyloseq object with the ASV table that has been filtered to remove low-abundance ASVs, no blanks
otu <- read.table("silva_nochloronomito_otu_table_bs.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_bs.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_bs.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps #456 taxa and 480 samples

# set metadata as factors for anosim/permanova
treat<-as.character(samples$treatment)
month<-as.character(samples$month)
health<-as.character(samples$condition)
tag<-as.character(samples$tag)
year<-as.character(samples$year)
site<-as.character(samples$site)

# permanova between groups using Aitchison distance
dist.clr <- dist(E.clr)
perm<-adonis(dist.clr~treat*month,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~treat*health,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~health*month,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~tag*month,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~tag*health,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~health*year,as(sample_data(ps),"data.frame"))
print(perm)
perm<-adonis(dist.clr~treat*site,as(sample_data(ps),"data.frame"))
print(perm)

# anosim between groups using Aitchison distance
dist.clr <- dist(E.clr)
ano.date <- anosim(dist.clr, date, permutations=999)
pdf("anosim_date.pdf",width=8.5)
plot(ano.date)
dev.off()
ano.treat <- anosim(dist.clr, treat, permutations=999)
pdf("anosim_treat.pdf",width=8.5)
plot(ano.treat)
dev.off()
ano.health <- anosim(dist.clr, health, permutations=999)
pdf("anosim_health.pdf",width=8.5)
plot(ano.health)
dev.off()

## Beta Diversity Dispersions
#calculate multivariate dispersions based on condition
mod <-betadisper(dist.clr, date)
#one way anova
anova(mod)
#boxplots
plot(mod)
boxplot(mod)

#calculate multivariate dispersions based on condition
mod2 <-betadisper(dist.clr, treat)
#one way anova
anova(mod2)
#boxplots
plot(mod2)
boxplot(mod2)


#calculate multivariate dispersions based on condition
mod3 <-betadisper(dist.clr, health)
#one way anova
anova(mod3)
#boxplots
plot(mod3)
pdf("distancetocentroid_health.pdf",width=8.5)
boxplot(mod3)
dev.off()
```



```{r, echo=FALSE}
otu <- read.table("silva_nochloronomito_otu_table_ps5.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps5.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_noblanks.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE),sample_data(samples),tax_table(taxon))
ps #456 taxa and 581 samples
bs3<-subset_samples(ps, site=="BS3")
bs3  #456 taxa and 221 samples
otu_bs3 = as(otu_table(bs3), "matrix")
taxon_bs3 = as(tax_table(bs3), "matrix")
meta_bs3 = as(sample_data(bs3), "matrix")
write.table(otu_bs3,"silva_nochloronomito_otu_table_bs3.txt",sep="\t",col.names=NA)
write.table(taxon_bs3,"silva_nochloronomito_taxa_table_bs3.txt",sep="\t",col.names=NA)
write.table(meta_bs3,"metadata_bs3.txt",sep="\t",col.names=NA)
otu <- read.table("silva_nochloronomito_otu_table_bs3.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_bs3.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_bs3.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE),sample_data(samples),tax_table(taxon))
ps #456 taxa and 221 samples
d.czm <- cmultRepl(t(otu), method="CZM", label=0)
d.clr <- codaSeq.clr(d.czm)
E.clr <- t(d.clr)
d.pcx <- prcomp(E.clr)
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)
# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
samples$month<-factor(samples$month,levels=c("July","Aug","Nov"))
samples$treatment <- factor(samples$treatment, levels = c("none","control paste and bag","probiotic paste and bag","resistant"))
treatcolors <- c("none"="#999999","control paste and bag"="#000066","probiotic paste and bag"="#993300","resistant"="#000000")

pdf("PCA_BS3.pdf",bg ="white",width=8.5)
p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=samples$treatment,shape=samples$month))
p<-p+geom_point(size=5)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  scale_fill_manual(values=treatcolors)+
  scale_shape_manual(values=c(21,22,24))+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab, fill="Treatment",shape="Month") + coord_fixed()
dev.off()
```





```{r, echo=FALSE}
##### bar charts
otu <- read.table("silva_nochloronomito_otu_table_ps5.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps5.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_noblanks.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps5 <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps5 #456 taxa and 581 samples
ps5_ra<-transform_sample_counts(ps5, function(OTU) OTU/sum(OTU))
ps5_ra #456 taxa and 581 samples
get_taxa_unique(ps5_ra, "Order") #67
get_taxa_unique(ps5_ra, "Class") #23
n <- 23
palette <- distinctColorPalette(n)
#you can rerun the previous line to get a new selection of colors
# keep list of colors used in palette that is most appealing
sink("palette.txt")
print(palette)
sink()

samples$month<-factor(samples$month,levels=c("Jan","May","July","Aug","Sept","Oct","Nov"))
samples$treatment <- factor(samples$treatment, levels = c("none","control bag","control paste","control paste and bag","probiotic bag","probiotic paste","probiotic paste and bag","resistant"))
treatcolors <- c("none"="#999999","control bag"="#56B4E9","control paste"="#0072B2","control paste and bag"="#000066","probiotic bag"="#E69F00","probiotic paste"="#D55E00","probiotic paste and bag"="#993300","resistant"="#000000")

pdf("barchart_Class.pdf",width=13)
p1=plot_bar(ps5_ra, fill="Class")+
  geom_bar(aes(fill=Class), stat="identity",position="stack")+
  facet_grid(.~treatment,scales="free",space="free")+
  theme(strip.text=element_text(face="bold"))+
  theme(axis.text.x=element_text(angle = 90))+
  scale_fill_manual(values=palette)+
  theme(plot.title = element_text(face="italic"))+
  theme(axis.title.x = element_blank())+
  theme(legend.position = "bottom")
p1
dev.off()
```

## Plot relative abundances of Pseudoalteromonas ASVs by treatment type


```{r, echo=FALSE}
bs3<-subset_samples(ps5_ra, site=="BS3")
bs3 #456 taxa and 221 samples
Pseudoalt<-subset_taxa(bs3, Genus=="Pseudoalteromonas")
Pseudoalt #6 taxa and 221 samples
otu.Pseudoalt = as(otu_table(Pseudoalt), "matrix")
otu.Pseudoalt<-as.data.frame(otu.Pseudoalt)
otu.Pseudoalt<-rownames_to_column(otu.Pseudoalt,var="Sample")
#export ASV sequences for supplemental table
write.table(otu.Pseudoalt,"Pseudoalteromonas_ASVs.txt",sep="\t",col.names=NA)

names(otu.Pseudoalt)[2]<-"ASV1"
names(otu.Pseudoalt)[3]<-"ASV2"
names(otu.Pseudoalt)[4]<-"ASV3"
names(otu.Pseudoalt)[5]<-"ASV4"
names(otu.Pseudoalt)[6]<-"ASV5"
names(otu.Pseudoalt)[7]<-"ASV6"

meta.Pseudoalt = as(sample_data(Pseudoalt), "matrix")
meta.Pseudoalt<-as.data.frame(meta.Pseudoalt)
meta.Pseudoalt<-rownames_to_column(meta.Pseudoalt,var="Sample")
otu.Pseudoalt.meta<-merge(meta.Pseudoalt,otu.Pseudoalt,"Sample")
otu_long_p<-reshape2::melt(otu.Pseudoalt.meta,id.vars=c("Sample","tag","treatment","site","species","condition","treated","month","year"),variable.name="ASV",value.name="Proportion")

#PaletteWoodsHole bog and rocky_beach
pseudoalt_colors<-c("ASV1"="#AC7070","ASV2"="#D6D9DA","ASV3"="#33291B","ASV4"="#C5C7C7","ASV5"="#937771","ASV6"="#5A4A35")

samples$treatment <- factor(samples$treatment, levels = c("none","control paste and bag","probiotic paste and bag","resistant"))
treatcolors <- c("none"="#999999","control paste and bag"="#000066","probiotic paste and bag"="#993300","resistant"="#000000")


# bubble plot, first need to detach plyr (or it overrides dplyr operations and the group_by doesn't work); detach Rmisc before plyr or it won't detach plyr
detach(package:Rmisc)
detach(package:plyr)
otu_long_p_ASV <- otu_long_p %>% group_by(treatment,ASV) %>% summarize(ASV_mean=mean(Proportion))
otu_long_p_ASV$treatment <- factor(otu_long_p_ASV$treatment, levels = c("none","control paste and bag","probiotic paste and bag","resistant"))

pdf("Pseudoalteromonas_ASVs_BS3.pdf",bg = "white",width=8,height=5)
p1<-ggplot(otu_long_p_ASV,aes(ASV,treatment,color=ASV))+
  geom_point(aes(size=ASV_mean))+
  theme_bw()+
  scale_size(range = c(0.5,10),name="Mean Relative Abundance")+
  scale_color_viridis(discrete=TRUE)+
  #scale_color_manual(values=pseudoalt_colors)+
  guides(color=FALSE)+  #removes the legend for the colors
  theme(axis.text.x=element_text(size=10,angle=90,hjust=1,vjust=0,color="black"))+
  theme(axis.text.y=element_text(size=12,color="black"))+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())
p1
dev.off()
```

## Plot all Pseudoalteromonas ASVs by Treatment

```{r, echo=FALSE}
otu_long_p$treatment <- factor(otu_long_p$treatment, levels = c("resistant","none","control paste and bag","probiotic paste and bag"))
otu_long_p$month<-factor(otu_long_p$month,levels=c("July","Aug","Nov"))
otu_long_p$condition<-factor(otu_long_p$condition,levels=c("HH","HD","DD"))
condcolors <- c("HH"="#E69F00", "HD"="#999999", "DD"="#000000")

#one-way anova
model1<- aov(otu_long_p$Proportion~otu_long_p$treatment)
summary(model1)
TukeyHSD(model1)

model2<- aov(otu_long_p$Proportion~otu_long_p$month)
summary(model2)
TukeyHSD(model2)

model3<- aov(otu_long_p$Proportion~otu_long_p$condition)
summary(model3)
TukeyHSD(model3)

pdf("Pseudoalteromonas_byTreatment_Date_BS3.pdf",width=8,height=6)
p2<-ggplot(otu_long_p,aes(x=treatment,y=Proportion))+
  geom_boxplot(outlier.shape=NA)+
  theme_bw()+
  facet_grid(month~.,space="free")+
  geom_jitter(position=position_jitter(width=.1, height=0),aes(color=condition),size=3)+
  scale_color_manual(values=condcolors)+
  theme(axis.title.x=element_blank())+
  theme(text=element_text(size=14))+
  theme(strip.text.y=element_text(size=14))+
  ylab("Relative Abundance of Pseudoalteromonas ASVs")
p2
dev.off()

pdf("Pseudoalteromonas_byTreatment.pdf",width=8,height=6)
p3<-ggplot(otu_long_p,aes(x=Treatment,y=Proportion))+
  geom_boxplot(outlier.shape=NA)+
  theme_bw()+
  #facet_grid(Date~.,space="free",labeller=labeller(Date = labels))+
  geom_jitter(position=position_jitter(width=.1, height=0),aes(color=HealthState),size=3)+
  scale_color_manual(values=condcolors)+
  theme(axis.title.x=element_blank())+
  theme(text=element_text(size=14))+
  theme(strip.text.y=element_text(size=14))+
  ylab("Relative Abundance of Pseudoalteromonas ASVs")
p3
dev.off()

# Repeat tests with only ASV1 (exact sequence match over 253-bp V4 region to Pseudoalteromonas strain McH1-7)
PseudoASV1 <- filter(otu_long_p, ASV == "ASV1")

#one-way anova
model1<- aov(PseudoASV1$Proportion~PseudoASV1$Treatment)
summary(model1)
TukeyHSD(model1)

model2<- aov(PseudoASV1$Proportion~PseudoASV1$Date)
summary(model2)
TukeyHSD(model2)

model3<- aov(PseudoASV1$Proportion~PseudoASV1$HealthState)
summary(model3)
TukeyHSD(model3)

```


## Plot relative abundances of Vibrio ASVs by treatment type

```{r, echo=FALSE}
vibrio<-subset_taxa(ps5_ra, Genus=="Vibrio")
vibrio
otu.vibrio = as(otu_table(vibrio), "matrix")
taxon.vibrio = as(tax_table(vibrio), "matrix")
meta.vibrio = as(sample_data(vibrio), "matrix")
otu.vibrio<-as.data.frame(otu.vibrio)
otu.vibrio<-rownames_to_column(otu.vibrio,var="Sample")
#export ASV sequences for supplemental table
write.table(otu.vibrio,"vibrio_ASVs.txt",sep="\t",col.names=NA)

names(otu.vibrio)[2]<-"ASV1"
names(otu.vibrio)[3]<-"ASV2"
names(otu.vibrio)[4]<-"ASV3"
names(otu.vibrio)[5]<-"ASV4"
names(otu.vibrio)[6]<-"ASV5"
names(otu.vibrio)[7]<-"ASV6"
names(otu.vibrio)[8]<-"ASV7"
names(otu.vibrio)[9]<-"ASV8"
names(otu.vibrio)[10]<-"ASV9"
names(otu.vibrio)[11]<-"ASV10"
names(otu.vibrio)[12]<-"ASV11"
names(otu.vibrio)[13]<-"ASV12"
names(otu.vibrio)[14]<-"ASV13"
names(otu.vibrio)[15]<-"ASV14"
names(otu.vibrio)[16]<-"ASV15"
names(otu.vibrio)[17]<-"ASV16"
names(otu.vibrio)[18]<-"ASV17"
names(otu.vibrio)[19]<-"ASV18"
names(otu.vibrio)[20]<-"ASV19"
names(otu.vibrio)[21]<-"ASV20"
names(otu.vibrio)[22]<-"ASV21"
names(otu.vibrio)[23]<-"ASV22"
names(otu.vibrio)[24]<-"ASV23"
names(otu.vibrio)[25]<-"ASV24"
names(otu.vibrio)[26]<-"ASV25"
names(otu.vibrio)[27]<-"ASV26"
names(otu.vibrio)[28]<-"ASV27"
names(otu.vibrio)[29]<-"ASV28"

meta.vibrio<-as.data.frame(meta.vibrio)
meta.vibrio<-rownames_to_column(meta.vibrio,var="Sample")
otu.vibrio.meta<-merge(meta.vibrio,otu.vibrio,"Sample")
otu_long<-reshape2::melt(otu.vibrio.meta,id.vars=c("Sample","TagID","HealthState","Treatment","Treated","AllPresent","Status","Date"),variable.name="ASV",value.name="Proportion")

# bubble plot, first need to detach plyr (or it overrides dplyr operations and the group_by doesn't work); detach Rmisc before plyr or it won't detach plyr
detach(package:Rmisc)
detach(package:plyr)
otu_long_ASV <- otu_long %>% group_by(Treatment,ASV) %>% summarize(ASV_mean=mean(Proportion))
#otu_long_ASV$radius <- sqrt(otu_long_ASV$ASV_mean / pi )
otu_long_ASV$Treatment <- factor(otu_long_ASV$Treatment, levels = c("none","control bag","control paste", "probiotic bag","probiotic paste"))

pdf("Vibrio_ASVs.pdf",bg = "white",width=10,height=5)
p1<-ggplot(otu_long_ASV,aes(ASV,Treatment,color=ASV))+
  geom_point(aes(size=ASV_mean))+
  scale_size(range = c(0.5,10),name="Mean Relative Abundance")+
  guides(color=FALSE)+  #removes the legend for the colors
  scale_color_viridis(discrete=TRUE)+
  theme(axis.text.x=element_text(size=10,angle=90,hjust=1,vjust=0,color="black"))+
  theme(axis.text.y=element_text(size=12,color="black"))+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())
p1
dev.off()

```

## Indicator species analysis

Add metadata columns to ASV table with relative abundances as shown at the top of this tutorial: https://jkzorz.github.io/2019/07/02/Indicator-species-analysis.html


```{r, echo=FALSE}

asv <- read.table("silva_nochloronomito_otu_table_ps5_RA_ForIndicspecies.txt",sep="\t",header=TRUE)
abund = asv[,6:ncol(asv)]
treated=asv$Treated
inv = multipatt(abund, treated, func = "r.g", control = how(nperm=9999))
summary(inv)

health=asv$HealthState
inv2 = multipatt(abund, health, func = "r.g", control = how(nperm=9999))
summary(inv2)

```

#ANCOM Function - compare across multiple treatments groups using a compositional appproach
#https://sites.google.com/site/siddharthamandal1985/research

#Need to run this first in order to run ANCOM on data

```{r, echo=FALSE}
ancom.W = function(otu_data,var_data,
                   adjusted,repeated,
                   main.var,adj.formula,
                   repeat.var,long,rand.formula,
                   multcorr,sig){

  n_otu=dim(otu_data)[2]-1

  otu_ids=colnames(otu_data)[-1]

  if(repeated==F){
    data_comp=data.frame(merge(otu_data,var_data,by="Sample.ID",all.y=T),row.names=NULL)
    #data_comp=data.frame(merge(otu_data,var_data[,c("Sample.ID",main.var)],by="Sample.ID",all.y=T),row.names=NULL)
  }else if(repeated==T){
    data_comp=data.frame(merge(otu_data,var_data,by="Sample.ID"),row.names=NULL)
    # data_comp=data.frame(merge(otu_data,var_data[,c("Sample.ID",main.var,repeat.var)],by="Sample.ID"),row.names=NULL)
  }

  base.formula = paste0("lr ~ ",main.var)
  if(repeated==T){
    repeat.formula = paste0(base.formula," | ", repeat.var)
  }
  if(adjusted==T){
    adjusted.formula = paste0(base.formula," + ", adj.formula)
  }

  if( adjusted == F & repeated == F ){
    fformula  <- formula(base.formula)
  } else if( adjusted == F & repeated == T & long == T ){
    fformula  <- formula(base.formula)
  }else if( adjusted == F & repeated == T & long == F ){
    fformula  <- formula(repeat.formula)
  }else if( adjusted == T & repeated == F  ){
    fformula  <- formula(adjusted.formula)
  }else if( adjusted == T & repeated == T  ){
    fformula  <- formula(adjusted.formula)
  }else{
    stop("Problem with data. Dataset should contain OTU abundances, groups,
         and optionally an ID for repeated measures.")
  }


  if( repeated==FALSE & adjusted == FALSE){
    if( length(unique(data_comp[,which(colnames(data_comp)==main.var)]))==2 ){
      tfun <- exactRankTests::wilcox.exact
    } else{
      tfun <- stats::kruskal.test
    }
  }else if( repeated==FALSE & adjusted == TRUE){
    tfun <- stats::aov
  }else if( repeated== TRUE & adjusted == FALSE & long == FALSE){
    tfun <- stats::friedman.test
  }else if( repeated== TRUE & adjusted == FALSE & long == TRUE){
    tfun <- nlme::lme
  }else if( repeated== TRUE & adjusted == TRUE){
    tfun <- nlme::lme
  }

  logratio.mat <- matrix(NA, nrow=n_otu, ncol=n_otu)
  for(ii in 1:(n_otu-1)){
    for(jj in (ii+1):n_otu){
      data.pair <- data_comp[,which(colnames(data_comp)%in%otu_ids[c(ii,jj)])]
      lr <- log((1+as.numeric(data.pair[,1]))/(1+as.numeric(data.pair[,2])))

      lr_dat <- data.frame( lr=lr, data_comp,row.names=NULL )

      if(adjusted==FALSE&repeated==FALSE){  ## Wilcox, Kruskal Wallis
        logratio.mat[ii,jj] <- tfun( formula=fformula, data = lr_dat)$p.value
      }else if(adjusted==FALSE&repeated==TRUE&long==FALSE){ ## Friedman's
        logratio.mat[ii,jj] <- tfun( formula=fformula, data = lr_dat)$p.value
      }else if(adjusted==TRUE&repeated==FALSE){ ## ANOVA
        model=tfun(formula=fformula, data = lr_dat,na.action=na.omit)
        picker=which(gsub(" ","",row.names(summary(model)[[1]]))==main.var)
        logratio.mat[ii,jj] <- summary(model)[[1]][["Pr(>F)"]][picker]
      }else if(repeated==TRUE&long==TRUE){ ## GEE
        model=tfun(fixed=fformula,data = lr_dat,
                   random = formula(rand.formula),
                   correlation=corAR1(),
                   na.action=na.omit)
        picker=which(gsub(" ","",row.names(anova(model)))==main.var)
        logratio.mat[ii,jj] <- anova(model)[["p-value"]][picker]
      }
    }
  }

  ind <- lower.tri(logratio.mat)
  logratio.mat[ind] <- t(logratio.mat)[ind]

  logratio.mat[which(is.finite(logratio.mat)==FALSE)] <- 1

  mc.pval <- t(apply(logratio.mat,1,function(x){
    s <- p.adjust(x, method = "BH")
    return(s)
  }))

  a <- logratio.mat[upper.tri(logratio.mat,diag=FALSE)==TRUE]

  b <- matrix(0,ncol=n_otu,nrow=n_otu)
  b[upper.tri(b)==T] <- p.adjust(a, method = "BH")
  diag(b)  <- NA
  ind.1    <- lower.tri(b)
  b[ind.1] <- t(b)[ind.1]

  #########################################
  ### Conservative
  if(multcorr==1){
    W <- apply(b,1,function(x){
      subp <- length(which(x<sig))
    })
    ### Moderate
  } else if(multcorr==2){
    W <- apply(mc.pval,1,function(x){
      subp <- length(which(x<sig))
    })
    ### No correction
  } else if(multcorr==3){
    W <- apply(logratio.mat,1,function(x){
      subp <- length(which(x<sig))
    })
  }

  return(W)
  }


ANCOM.main = function(OTUdat,Vardat,
                      adjusted,repeated,
                      main.var,adj.formula,
                      repeat.var,longitudinal,
                      random.formula,
                      multcorr,sig,
                      prev.cut){

  p.zeroes=apply(OTUdat[,-1],2,function(x){
    s=length(which(x==0))/length(x)
  })

  OTUdat.thinned=OTUdat[,c(1,1+which(p.zeroes<prev.cut))]

  otu.names=colnames(OTUdat.thinned)[-1]

  W.detected   <- ancom.W(OTUdat.thinned,Vardat,
                          adjusted,repeated,
                          main.var,adj.formula,
                          repeat.var,longitudinal,random.formula,
                          multcorr,sig)

  W_stat       <- W.detected

  W_frame = data.frame(otu.names,W_stat,row.names=NULL)
  W_frame = W_frame[order(-W_frame$W_stat),]

  W_frame$detected_0.9=rep(FALSE,dim(W_frame)[1])
  W_frame$detected_0.8=rep(FALSE,dim(W_frame)[1])
  W_frame$detected_0.7=rep(FALSE,dim(W_frame)[1])
  W_frame$detected_0.6=rep(FALSE,dim(W_frame)[1])

  W_frame$detected_0.9[which(W_frame$W_stat>0.9*(dim(OTUdat.thinned[,-1])[2]-1))]=TRUE
  W_frame$detected_0.8[which(W_frame$W_stat>0.8*(dim(OTUdat.thinned[,-1])[2]-1))]=TRUE
  W_frame$detected_0.7[which(W_frame$W_stat>0.7*(dim(OTUdat.thinned[,-1])[2]-1))]=TRUE
  W_frame$detected_0.6[which(W_frame$W_stat>0.6*(dim(OTUdat.thinned[,-1])[2]-1))]=TRUE

  final_results=list(W_frame)
  names(final_results)=c("W.taxa")
  return(final_results)
}
```

Perform ANCOM analysis

```{r, echo=FALSE}

#Read in count data, not relative abundance, no low-abundance ASVs removed
otu <- read.table("silva_nochloronomito_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps #20031 taxa and 196 samples
#Remove blanks
ps = subset_samples(ps, TagID != "blank")
ps #20031 taxa and 194 samples

get_taxa_unique(ps, "Family") #678
dat <- tax_glom(ps, taxrank = "Family")
datm <- psmelt(dat)
otud <- reshape2::dcast(datm, Sample ~ Family, value.var = 'Abundance', fun.aggregate = sum)
colnames(otud)[1] <- "Sample.ID"
metadat <- sample_data(ps) 
metadat <- as.data.frame(as.matrix(metadat))
metadat <- tibble::rownames_to_column(metadat, "Sample.ID")
names(otud) <- make.names(names(otud))
otu_test <- otud
write.table(otu_test,"ANCOM_otu_test.txt",sep="\t",col.names=NA)
metadat <- select(metadat, c("Sample.ID","HealthState","Treatment","Status","Date","TagID","Treated")) # use select to only use treatment columns of interest
map_test <- metadat
write.table(map_test,"ANCOM_map_test.txt",sep="\t",col.names=NA)

#bring in saved otu_test and map_test, if needed
otu_test<-read.table("ANCOM_otu_test.txt",sep="\t",header=T,row.names=1)
map_test<-read.table("ANCOM_map_test.txt",sep="\t",header=T,row.names=1)

Vardat <- map_test

#### ANCOM test - Treatment, not adjusted, more than 2 levels = Kruskal Wallis
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=FALSE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Treatment", #main variable or fator
                                 adj.formula= NULL, #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res1 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res1,"ANCOM_family_KruskallWallis_Treatment.txt",sep="\t",col.names=NA)
res2 <- res1[which(res1$detected_0.7==TRUE),] 

#### ANCOM test - Treatment, Adjusted by HealthState, ANOVA
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=TRUE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Treatment", #main variable or fator
                                 adj.formula= "HealthState", #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res3 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res3,"ANCOM_family_ANOVA_Treatment_adjHealthState.txt",sep="\t",col.names=NA)
res4 <- res3[which(res3$detected_0.7==TRUE),] 

#### ANCOM test - HealthState, no adjustment, KruskallWallis
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=FALSE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="HealthState", #main variable or fator
                                 adj.formula= NULL, #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res5 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res5,"ANCOM_family_KruskallWallis_HealthState.txt",sep="\t",col.names=NA)
res6 <- res5[which(res5$detected_0.7==TRUE),] 

#### ANCOM test - HealthState, Adjusted by Treatment, ANOVA
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=TRUE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="HealthState", #main variable or fator
                                 adj.formula= "Treatment", #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res7 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res7,"ANCOM_family_ANOVA_HealthState_adjTreatment.txt",sep="\t",col.names=NA)
res8 <- res7[which(res7$detected_0.7==TRUE),] 

#### ANCOM test - Status, no adjustment, KruskallWallis
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=FALSE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Status", #main variable or fator
                                 adj.formula= NULL, #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res9 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res9,"ANCOM_family_KruskallWallis_Status.txt",sep="\t",col.names=NA)
res10 <- res9[which(res9$detected_0.7==TRUE),] 

#### ANCOM test - Status, Adjusted by Treatment, ANOVA
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=TRUE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Status", #main variable or fator
                                 adj.formula= "Treatment", #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res11 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res11,"ANCOM_family_ANOVA_Status_adjTreatment.txt",sep="\t",col.names=NA)
res12 <- res11[which(res11$detected_0.7==TRUE),] 

#### ANCOM test - Date, no adjustment, KruskallWallis
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=FALSE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Date", #main variable or fator
                                 adj.formula= NULL, #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res13 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res13,"ANCOM_family_KruskallWallis_Date.txt",sep="\t",col.names=NA)
res14 <- res13[which(res13$detected_0.7==TRUE),] 

#### ANCOM test - Date, Adjusted by Treatment, ANOVA
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=TRUE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="Date", #main variable or fator
                                 adj.formula= "Treatment", #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res15 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res15,"ANCOM_family_ANOVA_Date_adjTreatment.txt",sep="\t",col.names=NA)
res16 <- res15[which(res15$detected_0.7==TRUE),]
res16b <- res15[which(res15$detected_0.9==TRUE),]

#### ANCOM test - Colony (TagID) no adjustment, KruskallWallis
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=FALSE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="TagID", #main variable or fator
                                 adj.formula= NULL, #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res17 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res17,"ANCOM_family_KruskallWallis_Colony.txt",sep="\t",col.names=NA)
res18 <- res17[which(res17$detected_0.7==TRUE),] 

#### ANCOM test - Colony, Adjusted by Date, ANOVA
comparison_test_treat=ANCOM.main(OTUdat=otu_test, #calling the OTU table
                                 Vardat=map_test, #calling the metadata
                                 adjusted=TRUE, #true if covariates are to be included for adjustment
                                 repeated=FALSE, #repeated measure
                                 main.var="TagID", #main variable or fator
                                 adj.formula= "Date", #other factors to include
                                 repeat.var=FALSE, #repeated measure
                                 long = FALSE, #longitudinal study
                                 multcorr=2,
                                 sig=0.05, #significance level
                                 prev.cut=0.90) #OTUs with proportion of zeroes greater than prev.cut are not included in the analysis
res19 <- comparison_test_treat$W.taxa #taxa that significantly vary across factor level of interest
write.table(res19,"ANCOM_family_ANOVA_Colony_adjDate.txt",sep="\t",col.names=NA)
res20 <- res19[which(res19$detected_0.7==TRUE),] 

```

Now plot the ANCOM results

```{r, echo=FALSE}

sig_sites <- glue::glue_collapse(droplevels(factor(res16b$otu.names)), sep = ", ") #this is to get a list of the families that are different
print(sig_sites)

#Calculate relative abundance
dim(otu_test) #194 rows 679 columns
otu_test_relabund <-  sweep(otu_test[,2:679], 1, rowSums(otu_test[,2:679]), '/')
otu_test_relnames <- cbind(map_test[,1:7],otu_test_relabund)

# Select only the significant differentially abundant families listed by sig_sites
# Note that any family names with punctuation must be enclosed with quotation marks!
sig_dis <- select(otu_test_relnames, "Sample.ID", HealthState, Treatment, Status, Date, TagID, Treated, Desulfovibrionaceae, DEV007, Rubritaleaceae, "Clostridiaceae_4", Clostridiales, FGL7S, Micrococcaceae, Halomonadaceae, Sedimenticolaceae, Thiovulaceae, Crocinitomicaceae, Microtrichaceae, Midichloriaceae, Rhodobacteraceae, Shewanellaceae, Cyanobiaceae, Saccharospirillaceae, Terasakiellaceae, Thiohalorhabdaceae, Vibrionaceae, Microbacteriaceae, Pseudoalteromonadaceae, Alphaproteobacteria, Geodermatophilaceae, Phormidiaceae, Flavobacteriaceae, Nitrincolaceae, Phycisphaeraceae, Bacteria, Desulfobulbaceae, Thiomicrospiraceae, Rhizobiaceae, Rhodanobacteraceae)
sig_long <- reshape2::melt(sig_dis, id.vars=c("Sample.ID","HealthState","Treatment","Status","Date","TagID","Treated"),variable.name="Family",value.name="Proportion")
sum_sig <- Rmisc::summarySE(sig_long, measurevar = "Proportion", groupvars = c("Date","Family"), na.rm=TRUE)

sum_sig$Date<-factor(sum_sig$Date,levels=c("A","O","J"))
cols <- c("A"="#56B4E9", "O"="#E69F00", "J"="#999999")

pdf("ANCOM_Date_adjTreatment.pdf")
fams <- ggplot(sum_sig, aes(x=Family, y=Proportion))+
  geom_point(size=4, aes(color=Date))+
  scale_color_manual(values=cols)+
  coord_flip()+
  theme_bw()+
  theme(axis.text.x=element_text(size=14))+
  theme(axis.text.y=element_text(size=14))+
  theme(axis.title.x=element_text(size=14))+
  theme(axis.title.y=element_text(size=14))+
  #theme(legend.justification=c(1,1), legend.position=c(1,1))+
  geom_errorbar(aes(ymin=Proportion-se, ymax=Proportion+se), width=.1)+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size=12))
fams
dev.off()
```

